---
title: "PA1_template"
author: "DanLarson"
date: "February 13, 2016"
output: html_document
---

## Part 1

Load the data and the required packages. For this assignment I used ggplot2, dplyr, and tidyr. ggplot2 is for the figure development whil dplyr and tidyr are used for data wrangling. 

```{r,echo=FALSE}

# load required packages
require(ggplot2)
require(dplyr)
require(tidyr)

activity <- tbl_df(read.csv("activity.csv",header=TRUE))


```

##Part 2

Develop a histogram that shows the distribution of total steps per day. 

```{r}
require(ggplot2)
require(dplyr)
require(tidyr)
activity$date <- as.Date(activity$date)
activity <- subset(activity,!is.na(steps))
activityDay <- group_by(activity,date) %>% summarise(total_steps = sum(steps))
ggplot(data=activityDay,aes(x=total_steps))+geom_histogram(bins = 10,col="red")+ggtitle("Distribution of Total Steps per Day")
```

##Part 3

Mean and median number of steps taken each day

```{r}
summary(activity$steps)
```

## Part 4

Time series plot of the average number of steps taken

```{r}
require(ggplot2)
require(dplyr)
require(tidyr)
activityInt <- group_by(activity,interval) %>% summarise(total_steps = mean(steps))
ggplot(data=activityInt,aes(x=interval,y=total_steps))+geom_line()+
  ggtitle("Mean Steps per 5 Minute Interval")
```

## Part 5

The 5-minute interval that, on average, contains the maximum number of steps


```{r}
require(ggplot2)
require(dplyr)
require(tidyr)
head(arrange(activityInt,desc(total_steps)))
```


## Part 6

Code to describe and show a strategy for imputing missing data

```{r,echo=FALSE}

# load required packages
require(ggplot2)
require(dplyr)
require(tidyr)

activityNew <- activity
nas <- is.na(activityNew$steps)
avg_interval <- tapply(activityNew$steps, activityNew$interval, mean, na.rm=TRUE, simplify=TRUE)
activityNew$steps[nas] <- avg_interval[as.character(activityNew$interval[nas])]

```

## Part 7 

Histogram of the total number of steps taken each day after missing values are imputed

```{r}

# load required packages
require(ggplot2)
require(dplyr)
require(tidyr)

activityNew <- subset(activityNew,!is.na(steps))
activityDay <- group_by(activityNew,date) %>% summarise(total_steps = sum(steps))
ggplot(data=activityDay,aes(x=total_steps))+geom_histogram(bins = 10,col="red")+
  ggtitle("Distribution of Steps with Missing data replaced")


```

## Part 8

Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r}

# load required packages
require(ggplot2)
require(dplyr)
require(tidyr)

# Part 8
#Identify Weekday
activityNew$dow<-format(activityNew$date,
                     format = "%A")
#Identify Weekend/Weekday feature
activityNew$weekEnd <- ifelse(activityNew$dow=="Saturday","Weekend",
                              ifelse(activityNew$dow=="Sunday","Weekend","Weekday"))

#Panel Plot
activityIntWeekend <- group_by(activityNew,weekEnd,interval) %>% summarise(total_steps = mean(steps))
d<-ggplot(data=activityIntWeekend,aes(x=interval,y=total_steps,color=weekEnd))+
            geom_line()+
            facet_wrap(~weekEnd,nrow = 2,ncol = 1)+
            ggtitle("Mean Steps per 5 Minute Interval")
ggsave(d,file="Figures/stepIntervalWeekend.png",height = 6,width = 6)



```


